A Large-Area, Spatially Continuous Assessment of Land Cover Map Error and Its Impact on Downstream Analyses Relevant to Global Change

### 2017-03-22 paper/main.tex
Some more edits to Discussion and new title

\title{How Landcover Errors Impact Our Understanding of Socioeconomic and Environment Change}
\author[1,2,3]{Lyndon Estes}
\author[4]{Peng Chen}
\author[2]{Stephanie Debats}
\author[4]{Tom Evans}
\author[5]{Stefanus Ferreira}
\author[6]{Tobias Kuemmerle}
\author[2]{Gabrielle Ragazzo}
\author[2,7]{Justin Sheffield}
\author[8]{Adam Wolf}
\author[2]{Eric Wood}
\author[2,9]{Kelly Caylor}

@@

%\begin{article}
\begin{abstract}
{Land cover maps increasingly underlie research into socioeconomic and environmental processes, including global change. It is known that map errors impact our understanding of these processes, but quantifying these impacts is difficult because many areas lack adequate reference data. We used a high accuracy map of South African cropland to assess 1) the magnitude of error in several current generation land cover maps, and 2) how these errors propagate in downstream studies. We first quantified pixel-wise errors in the cropland classes of four widely used land cover maps at resolutions ranging from 1-100 km, then calculated errors in several representative ``downstream'' (map-based) analyses, including estimates of vegetative carbon stocks and crop production, simulated evapotranspiration, and agent-based assessment of household food security. We also assessed maps' spatial accuracy based on how precisely they could be used to locate specific landscape features. We found that cropland maps can have substantial biases and poor accuracy at all resolutions (e.g. at 1 km resolution, up to $\sim$45\% underestimates of cropland (bias) and nearly 50\% mean absolute error (accuracy); at 100 km, up to 15\% underestimates and nearly 20\% MAE). National-scale maps derived from higher resolution imagery were most accurate, followed by multi-map fusion products. Constraining mapped values to match survey statistics proved effective at minimizing bias (provided the statistics are accurate). Downstream errors could be substantially amplified or muted, depending on the values ascribed to cropland-adjacent covers (e.g. with forest as adjacent cover, carbon map error was 200-500\% greater than in the foundational cropland maps, but $\sim$40\% less for sparse cover types). The average locational error was 6 km (600\%). These findings provide deeper insight into the causes and consequences of land cover map error, including their potential broader regional and policy consequences, and suggest several recommendations for land cover map users.  
%
%
%These maps can have substantial error that impact , which may impact our understanding of environmental change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions. We used a high quality, national-scale cropland map to assess bias and accuracy within the cropland classes of current generation land cover datasets and four examples of ``downstream'' (land cover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km. We found that cropland maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors. In some cases these error metrics remain high even when maps are aggregated to 100 km. These errors can be substantially amplified in downstream analyses (e.g. the carbon and crop production analyses), but in cases where the values of different cover types that underpin the analysis are similar (e.g. evapotranspiration estimates) the results can be relatively insensitive. To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of land cover maps from their base resolution is often needed. NEW TEXT HERE.  
}
\end{abstract}

@@

The value of the insights resulting from such studies depends upon the veracity of their underlying land cover data \citep{verburg_challenges_2011}, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that this house has shaky foundations. The reason for this is that land cover data can only practically be derived from satellite imaging, which has several important constraints that propagate mapping errors. First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g. smallholders' farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}. Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g. croplands or savannas;][]{fritz_identifying_2008,verburg_challenges_2011,debats_generalized_2016}) and thus be difficult for the classifier to distinguish. Discretizing a continuous cover type (e.g. dividing a forest into different canopy cover classes) can promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}. Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013,fritz_identifying_2008}. Third, land cover maps are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, but seasonal variability and land cover changes can be easily confused. Given these multiple sources of error, land cover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011,fritz_need_2013}. These errors limit our ability to obtain granular, mechanistic understanding of processes related to global change.

@@

Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}. Although a number of previous studies have examined how land cover maps errors can propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas. \citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014,schmit_limitations_2006}.

@@

We used these vector data as a reference for evaluating four land cover products representative of the type commonly used in global change studies and related areas of research. The first was South Africa's own 30 m resolution 2009 National land cover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}. Although global-scale, Landsat-derived maps have recently become available \citep{chen_global_2015}, their reported accuracy for cultivated areas is lower (80-85\% user's accuracy) than that achieved by more intensive, national to sub-national products \citep[e.g. 90\% user's accuracy][]{sweeney_mapping_2015}. The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover \citep{friedl_modis_2010} data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}. The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics. This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover. Since GeoWiki provides a continuous bounded value of land cover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.

@@

\vspace{-0.3 cm}
\subsection*{Recommendations}
\vspace{-0.2 cm}
These findings suggest several recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}. First, to minimize error, users should typically prefer maps derived from imagery with resolutions finer than the scale of individual objects in the cover(s) of interest, assuming that they were created with rigorous classification methods, report appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}. Finer resolution not only helps to improve classification accuracy, but can minimize the \emph{aggregation problem}, one of two components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence the results of analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}. In remote sensing, an image's pixels define the fundamental mapping unit (by summarizing reflectance or radar/lidar sensors), and the pixels' dimensions influence analyses as a function of how closely they match the characteristic shapes and scales of natural features \citep{dark_modifiable_2007}. However, if the sensor resolution is fine enough, then mismatches can be minimized--a natural feature's shape can be approximated by aggregating several square pixels--giving the analyst greater ability to minimize errors associated with this aspect of MAUP \citep{dark_modifiable_2007, hay_comparison_2003}.

Of course, high quality, fine-scaled maps such as SA-LC, which was carefully developed for South Africa, do not exist for many countries.  In the absence of these, the new 30 m GLOBELAND30 \citep{chen_global_2015} provides global coverage for the periods 2000 and 2010, and may prove useful in smaller regions lacking their own focused maps. Users should first conduct their own accuracy assessments within their regions of interest to develop an appropriate understanding of their error. It is also becoming increasingly feasible for researchers to make their own high resolution maps, due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g. cropland) and may be exacerbated by greater image resolution \citep[e.g.][]{debats_generalized_2016,sweeney_mapping_2015}.

If existing high-resolution maps are not available or cannot be readily made, users interested in agricultural land cover may choose to use a new fusion map, such as Geowiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}. Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error.  By calibrating against inventory data, these maps can also reduce pixel level biases (e.g. Fig. 1B). However, this adjustment technique surely introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries. \citep{carletto_emperor_2013,carletto_guesstimates_2015}.

Lastly, if a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover types, then this may be achieved by aggregating maps expressing continuous values (e.g. fractional cover) to coarser scales. This approach must be undertaken with care, as its poses a number of complications related to the \emph{scale problem}, the other half of the MAUP \citep{openshaw_million_1979}. These include progressive decline in variance with increasing scale \citep[even if means remain constant;][]{dark_modifiable_2007}, and reduced efficiency in estimating regression parameters from coarser map values \citep[][]{avelino_goldilocks_2016}. Nevertheless, aggregating continuous variables can reduce pixel errors without biasing some statistical properties (e.g. total cropland area remains constant using the aggregation approach demonstrated here; see Results: Cropland map errors), and has been shown to reduce other MAUP-related analytical problems \citep{avelino_goldilocks_2016}. However, the user must ensure that the scale of aggregation is appropriate for the analysis at hand.

>>>>

%\title{A Large-Area, Spatially Continuous Quantification of Error in Land Cover Maps and Associated Socioeconomic and Environmental Analyses Relevant to Global Change}
\title{A Large-Area, Spatially Continuous Assessment of Land Cover Map Error and Its Impact on Downstream Analyses Relevant to Global Change}
%\title{How Land Cover Errors Can Impact Understanding of Socioeconomic and Environment Phenomena}
\author[1,2,3]{Lyndon Estes}
\author[4]{Peng Chen}
\author[2]{Stephanie Debats}
\author[4]{Tom Evans}
\author[5]{Stefanus Ferreira}
\author[6]{Tobias Kuemmerle}
\author[2]{Gabrielle Ragazzo}
\author[2,7]{Justin Sheffield}
\author[8]{Adam Wolf}
\author[2]{Eric Wood}
\author[2,9]{Kelly Caylor}

@@

%\begin{article}
\begin{abstract}
{Land cover maps increasingly underlie research into socioeconomic and environmental patterns and processes, including global change. It is known that map errors impact our understanding of these phenomena, but quantifying these impacts is difficult because many areas lack adequate reference data. We used a high accuracy map of South African cropland to assess 1) the magnitude of error in several current generation land cover maps, and 2) how these errors propagate in downstream studies. We first quantified pixel-wise errors in the cropland classes of four widely used land cover maps at resolutions ranging from 1-100 km, then calculated errors in several representative ``downstream'' (map-based) analyses, including estimates of vegetative carbon stocks and crop production, simulated evapotranspiration, and agent-based assessment of household food security. We also assessed maps' spatial accuracy based on how precisely they could be used to locate specific landscape features. We found that cropland maps can have substantial biases and poor accuracy at all resolutions (e.g. at 1 km resolution, up to $\sim$45\% underestimates of cropland (bias) and nearly 50\% mean absolute error (accuracy); at 100 km, up to 15\% underestimates and nearly 20\% MAE). National-scale maps derived from higher resolution imagery were most accurate, followed by multi-map fusion products. Constraining mapped values to match survey statistics proved effective at minimizing bias (provided the statistics are accurate). Downstream errors could be substantially amplified or muted, depending on the values ascribed to cropland-adjacent covers (e.g. with forest as adjacent cover, carbon map error was 200-500\% greater than in the foundational cropland maps, but $\sim$40\% less for sparse cover types). The average locational error was 6 km (600\%). These findings provide deeper insight into the causes and consequences of land cover map error, including their potential broader regional and policy consequences, and suggest several recommendations for land cover map users.  
%
%
%These maps can have substantial error that impact , which may impact our understanding of environmental change processes and associated policies is thus of great interest, but is difficult to fully quantify because we lack spatially comprehensive reference data, particularly in the World's most rapidly developing regions. We used a high quality, national-scale cropland map to assess bias and accuracy within the cropland classes of current generation land cover datasets and four examples of ``downstream'' (land cover-dependent) analyses, two related to physical processes (estimates of vegetative carbon stocks and evapotranspiration) and two to socio-economic processes (gridded crop production estimates and agent-based simulation of household food security), at resolutions ranging from 1-100 km. We found that cropland maps have substantial errors below 25 km resolution, with substantial biases and inaccuracies, particularly from maps derived from coarse resolution sensors. In some cases these error metrics remain high even when maps are aggregated to 100 km. These errors can be substantially amplified in downstream analyses (e.g. the carbon and crop production analyses), but in cases where the values of different cover types that underpin the analysis are similar (e.g. evapotranspiration estimates) the results can be relatively insensitive. To avoid these errors, and thus minimize the risk of misunderstanding global change processes or misinforming policy, substantial map aggregation of land cover maps from their base resolution is often needed. NEW TEXT HERE.  
}
\end{abstract}

@@

The value of the insights resulting from such studies depends upon the veracity of their underlying land cover data \citep{verburg_challenges_2011}, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence indicates that this house has shaky foundations \citep[e.g.][]{fritz_highlighting_2011}. The reason for this is that land cover data can only practically be derived from satellite imaging, which has several important constraints that propagate mapping errors. First, in many regions the spatial arrangement of cover types is smaller than the sensor resolution \citep[e.g. smallholders' farms][]{jain_mapping_2013,debats_generalized_2016,ozdogan_resolution_2006}, or the types of interest are spectrally indistinct from neighboring ones \citep{sweeney_mapping_2015,fritz_identifying_2008}, which are factors that increase mapping complexity \citep{yu_meta-discoveries_2014}. Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties \citep[e.g. croplands or savannas;][]{fritz_identifying_2008,verburg_challenges_2011,debats_generalized_2016} and can thus be difficult for the classifier to distinguish. Discretizing a continuous cover type (e.g. dividing a forest into different canopy cover classes) can promote classification error, particularly near class boundaries \citep{foody_status_2002}, as well as confusion about the actual extent of the cover type \citep{sexton_conservation_2015}. Furthermore, class definitions often vary between maps, complicating inter-comparison \citep{kuemmerle_challenges_2013,fritz_identifying_2008}. Third, land cover maps are often used to detect changes \citep[e.g.][]{gross_monitoring_2013}, but seasonal variability and land cover changes can be easily confused. Given these multiple sources of error, land cover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \citep{estes_projected_2013,fritz_comparison_2010,fritz_cropland_2011,fritz_need_2013}. These errors limit our ability to obtain granular, mechanistic understanding of processes related to global change.

@@

Since it is difficult to fully quantify land cover map errors, it is even more challenging to gauge their impact on downstream analyses, where there is substantial risk of error amplification \citep{kuemmerle_challenges_2013}. Although a number of previous studies have examined how land cover maps errors can propagate, these are primarily assessed using either simulated errors, relative differences in existing land cover maps, or ground validation data covering relatively small areas \citep[e.g.][]{ge_impacts_2007, linard_assessing_2010, quaife_impact_2008,tuanmu_global_2014,schmit_limitations_2006}.

@@

We used these vector data as a reference for evaluating four land cover products representative of the type commonly used in global change studies and related areas of research. The first was South Africa's own 30 m resolution 2009 National land cover map \citep[SA-LC][]{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are created for individual countries \citep[e.g.][]{fry_completion_2009}. Although global-scale, Landsat-derived maps have recently become available \citep{chen_global_2015}, their reported accuracy for cultivated areas is lower (80-85\% user's accuracy) than that achieved by more intensive, national to sub-national products \citep[e.g. 90\% user's accuracy][]{sweeney_mapping_2015}. The second and third were respectively the 300 m GlobCover 2009 \citep{arino_global_2012} and 500 m resolution MODIS land cover \citep{friedl_modis_2010} data (for 2011, the final year of the reference interval), which are widely used global-scale products \citep[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}. The fourth dataset was the 1 km GeoWiki cropland map for Africa \citep{fritz_mapping_2015}, which fuses the cropland classes of multiple landcover products (including the three assessed here), with the resulting cropland percentages adjusted until their totals sum to match cropland areas reported in national agricultural statistics (this method also incorporated additional categorical training data collected via crowdsourcing). This ``hybrid-fusion'' approach represents the state of the art in mapping agricultural land cover. Since GeoWiki provides a continuous bounded value of land cover (cropland percentage), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.

@@

\vspace{-0.3 cm}
\subsection*{Recommendations}
\vspace{-0.2 cm}
These findings suggest a number of recommendations for using land cover maps, complementing those suggested in earlier work \citep{verburg_challenges_2011}, and beginning with those related to the selection of base land cover maps. First, to minimize error, users should typically prefer maps derived from imagery with resolutions finer than the scale of individual objects in the cover(s) of interest, assuming that they were created with rigorous classification methods accompanied by appropriate error metrics \citep{olofsson_good_2014}, and are thematically appropriate for the intended use \citep{verburg_challenges_2011}. Finer resolution not only helps to improve classification accuracy (see preceding section on \emph{Sources of error in cropland maps)}, but can minimize the \emph{aggregation problem}, one of two components of the modifiable area unit problem \citep[MAUP; ][]{openshaw_million_1979}, in which the shape and placement of the non-overlapping units used to extract map values influence the results of analyses of those values \citep{dark_modifiable_2007, marceau_scale_1999}. In remote sensing, an image's pixels define the fundamental mapping unit (by summarizing reflectance or radar/lidar sensors), and the pixels' dimensions influence analyses as a function of how closely they match the characteristic shapes and scales of natural features \citep{dark_modifiable_2007}. However, if the sensor resolution is fine enough, then mismatches can be minimized--a natural feature's shape can be approximated by aggregating several square pixels--giving the analyst greater ability to minimize errors associated with this aspect of MAUP \citep{dark_modifiable_2007, hay_comparison_2003}.

Of course, high quality, fine-scaled maps such as SA-LC, which was carefully developed for South Africa, do not exist for many countries.  In the absence of these, the new 30 m GLOBELAND30 \citep{chen_global_2015} provides global coverage for the periods 2000 and 2010, and may prove useful in smaller regions lacking their own focused maps. Users should first conduct their own, more-focussed accuracy assessments to better understand error rates within their region of interest. Another option is to make one's own high resolution maps; although harder to do, this is becoming increasingly feasible due to the rapidly expanding spatial and temporal coverage of high resolution imagery \citep[e.g.][]{hand_startup_2015,drusch_sentinel-2:_2012}, growing accessibility of high performance computing \citep[e.g.][]{pekel_high-resolution_2016}, and improvements to classification algorithms, particularly those designed to deal with the high intra-class variability inherent to certain cover types (e.g. cropland) and may be exacerbated by greater image resolution \citep[e.g.][]{debats_generalized_2016,sweeney_mapping_2015}.

If existing high-resolution maps are not available or cannot be readily made, users interested in agricultural land cover may choose to use a new fusion map, such as GeoWiki and its derivatives \citep{fritz_mapping_2015,waldner_unified_2016}. Although these maps are relatively coarse-scaled, the fusion methodology helps to greatly minimize error, as does the use of fractional cover values.  By calibrating against inventory data, these maps can also reduce pixel level biases (e.g. Fig. 1B). However, users should be aware that this adjustment technique introduces confounding errors where census-reported statistics are unreliable, such as in some SSA countries \citep{carletto_emperor_2013,carletto_guesstimates_2015}.

If a particularly erroneous map is all that is available, and the user is chiefly interested in minimizing cell-wise error and less concerned about the spatial configuration of cover types, then this may be achieved by aggregating maps expressing continuous values (e.g. fractional cover) to coarser scales. This approach must be undertaken with care, as its poses a number of complications related to the \emph{scale problem}, the other half of the MAUP \citep{openshaw_million_1979}. These include progressive decline in variance with increasing scale \citep[even if means remain constant;][]{dark_modifiable_2007}, and reduced efficiency in estimating regression parameters from coarser map values \citep[][]{avelino_goldilocks_2016}. Nevertheless, aggregating continuous variables can reduce pixel errors without biasing some statistical properties (e.g. total cropland area remains constant using the aggregation approach demonstrated here; see Results: Cropland map errors), and has been shown to reduce other MAUP-related analytical problems \citep{avelino_goldilocks_2016}. However, the user must ensure that the scale of aggregation is appropriate for the analysis at hand.

Finally, users (or makers) of downstream products should rigorously ascertain how error propagates from the base land cover map into their derived maps \citep{verburg_challenges_2011}. In some cases, the providers of ``off-the-shelf'' products report pixel-level uncertainty values \citep[e.g.][]{ramankutty_farming_2008}, which may be sufficient.  In other cases, downstream maps may lack quantified confidence intervals \citep[e.g.][]{monfreda_farming_2008}. Although such maps may provide guidelines for appropriate usage\footnote{www.earthstat.org\/wp-content\/uploads\/METADATA\_HarvestedAreaYield175Crops.pdf}, which our analyses help to further illustrate, users should undertake their own two-step error assessments. The most straightforward way may be to use a Monte Carlo approach that generates artificial datasets \citep[e.g.][]{avelino_goldilocks_2016} with introduced error, drawn from either reported or user-determined accuracy statistics, for both the base land cover and the downstream maps. Quantifying error propagation in this (or another appropriate) manner is important to understanding how map error may influence subsequent understanding of the phenomenon of interest.
